---
title: "Lab3A"
author: "Rovin Sharma"
date: "9/7/2020"
output: html_document
---

# Starting with the Data

```{r, message = FALSE, error = FALSE}
library(tidyverse)

download.file(url="https://ndownloader.figshare.com/files/2292169",
              destfile = "data/portal_data_joined.csv")

surveys <- read_csv("data/portal_data_joined.csv")              
```

**Challenge 1**

```{r}
 str(surveys)
``` 
The class of object "surveys" is data.frame, it has 34786 rows and 13 columns.
There are 48 species, but I am not sure how I could find that from the output of str.

**Challenge 2**

```{r}
surveys_200 <- surveys[200, ]

surveys_last <- surveys[nrow(surveys), ]

surveys_middle <- surveys[nrow(surveys)/2, ]

surveys[-(7:nrow(surveys)), ]
```

**Factors**

```{r}
sex <- factor(c("male", "female", "female", "male"))

levels(sex)
nlevels(sex)

sex <- factor(sex, levels = c("male", "female"))
sex
```

**Converting Factors**

```{r}
year_fct <- factor(c(1990, 1983, 1977, 1998, 1990))
as.numeric(as.character(year_fct))
as.numeric(levels(year_fct))[year_fct]
year_fct
```

**Renaming Factors**

```{r}
sex <- factor(surveys$sex)
head(sex)
levels(sex)

levels(sex)[1] <- "undetermined"
```

**Challenge 3**

```{r}
levels(sex)[2] <- "female"
levels(sex)[3] <- "male"

sex <- factor(sex, levels = c("female", "male", "undetermined"))
plot(sex)
```

**Challenge 4**

```{r}
animal_data <- data.frame(
          animal = c("dog", "cat", "sea cucumber", "sea urchin"),
          feel = c("furry", "squishy", "leathery", "spiny"),
          weight = c(45, 8, 1.1, 0.8)
          )
```

```{r}
country_climate <- data.frame(
       country = c("Canada", "Panama", "South Africa", "Australia"),
       climate = c("cold", "hot", "temperate", "hot/temperate"),
       temperature = c(10, 30, 18, 15),
       northern_hemisphere = c(TRUE, TRUE, FALSE, FALSE),
       has_kangaroo = c(FALSE, FALSE, FALSE, TRUE)
       )
```
The characters are treated as factors, but they would've been vectors with stringsAsFactors = FALSE. My code has the changes to reflect the right data type. The quotes in northern_hemisphere and temperature were removed, and the 1 in has_kangaroo was changed to TRUE.

*Formatting Dates*

```{r}

library("lubridate")

my_date <- ymd("2015-01-01")

my_date <- ymd(paste("2015", "1", "1", sep = "-")) 


paste(surveys$year, surveys$month, surveys$day, sep = "-")
ymd(paste(surveys$year, surveys$month, surveys$day, sep = "-"))

surveys$date <- ymd(paste(surveys$year, surveys$month, surveys$day, sep = "-"))

summary(surveys$date)

missing_dates <- surveys [is.na(surveys$date), c("year", "month", "day")]
head(missing_dates)
```
