---
title: "Lab 4"
author: "Rovin Sharma"
date: "9/20/2020"
output: html_document
---

#Setup
```{r, message = FALSE, erro = FALSE}
library(tidyverse)
library(DT)

report_03_11_2020 <-   read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/03-11-2020.csv")) %>%
  rename(Country_Region = "Country/Region", Province_State = "Province/State")

str(report_03_11_2020)  
view(report_03_11_2020)
```

#DT
```{r}
datatable(report_03_11_2020)
```


#Exercises Part 1

*1.*
```{r}
report_03_12 <- read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/03-12-2020.csv")) %>% 
  rename(Country_Region = "Country/Region", Province_State = "Province/State")
  
report_09_12 <- read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/09-12-2020.csv"))

view(report_03_12)
str(report_03_12)

view(report_09_12)
str(report_09_12)
```
The changes between 3/12/2020 and 9/12/2020 data tables:
3/12 has fewer rows as expected, but it also has fewer columns. 8 columns compared to the 14 columns in 9/12.
9/12 has changed some column names, simplifying latitude and longitude to Lat and Long, respectively.
9/12 also replaced "/" with "_"

*2.*
```{r}
report_03_12 %>% 
  group_by(Country_Region) %>% 
  summarize(States_Provinces = n(), 
            total_confirmed = sum(Confirmed)) %>% 
            datatable()


report_09_12 %>% 
  group_by(Country_Region) %>% 
  summarize(States_Provinces = n(),
            total_confirmed = sum(Confirmed)) %>% 
            datatable()
```
*3.*
```{r}
report_03_12 %>% 
  filter(Country_Region == "US") %>% 
  group_by(Country_Region) %>% 
  summarize(States = n(),
            total_confirmed = sum(Confirmed))
            
report_09_12 %>% 
  filter(Country_Region == "US") %>% 
  group_by(Country_Region) %>% 
  summarize(total_confirmed = sum(Confirmed))
```
*4.*
```{r}
report_03_12 %>% 
  group_by(Country_Region) %>% 
  summarize(total_deaths = sum(Deaths)) %>% 
  arrange(desc(total_deaths)) %>% 
  slice_head(n = 10) %>% 
  datatable()
 
report_09_12 %>% 
  group_by(Country_Region) %>% 
  summarize(total_deaths = sum(Deaths)) %>% 
  arrange(desc(total_deaths)) %>% 
  slice_head(n = 10) %>% 
  datatable()
```
*5.*
Setting up objects sorted with necessary data for plots
```{r}
report_03_12_cases <- report_03_12 %>% 
                      group_by(Country_Region) %>% 
                      summarize(total_cases = sum(Confirmed)) %>% 
                      arrange(desc(total_cases)) %>% 
                      slice_head(n = 10)

report_03_12_deaths <- report_03_12 %>% 
                      group_by(Country_Region) %>% 
                      summarize(total_deaths = sum(Deaths)) %>% 
                      arrange(desc(total_deaths)) %>% 
                      slice_head(n = 10)
                      
report_09_12_cases <- report_09_12 %>% 
                      group_by(Country_Region) %>% 
                      summarize(total_cases = sum(Confirmed)) %>% 
                      arrange(desc(total_cases)) %>% 
                      slice_head(n = 10)  

report_09_12_deaths <- report_09_12 %>% 
                      group_by(Country_Region) %>% 
                      summarize(total_deaths = sum(Deaths)) %>% 
                      arrange(desc(total_deaths)) %>% 
                      slice_head(n = 10)
```                      
Creating the plots
```{r}
report_03_12_cases_plot <- ggplot(report_03_12_cases, aes(x = Country_Region, y = total_cases)) + geom_point()

report_03_12_deaths_plot <- ggplot(report_03_12_deaths, aes(x = Country_Region, y = total_deaths)) + geom_point()

report_09_12_cases_plot <- ggplot(report_09_12_cases, aes(x = Country_Region, y = total_cases)) + geom_point()

report_09_12_deaths_plot <- ggplot(report_09_12_deaths, aes(x = Country_Region, y = total_deaths)) + geom_point()
```
*6.*
Creating the bar plots
```{r}
report_03_12_cases_bar <- ggplot(report_03_12_cases, aes(x = Country_Region, y = total_cases)) + geom_bar(stat = 'identity')

report_03_12_deaths_bar <- ggplot(report_03_12_deaths, aes(x = Country_Region, y = total_deaths)) + geom_bar(stat = 'identity')

report_09_12_cases_bar <- ggplot(report_09_12_cases, aes(x = Country_Region, y = total_cases)) + geom_bar(stat = 'identity')

report_09_12_deaths_bar <- ggplot(report_09_12_deaths, aes(x = Country_Region, y = total_deaths)) + geom_bar(stat = 'identity')
```
                      