---
title: "Lab 4a"
author: "Rovin Sharma"
date: "9/20/2020"
output: html_document
---

This is the auxiliary part of Lab 4 that contains examples and challenges from "Data visualization with ggplot2"

# Setup

```{r, include = FALSE, message = FALSE, error = FALSE}
library(tidyverse)
```
```{r}
surveys_complete <- read_csv("data/surveys_complete.csv")
```

# Building ggplot

```{r}
ggplot(data = surveys_complete)
 
ggplot(data = surveys_complete, mapping = aes(x = weight, y = hindfoot_length))
```

*adding 'geoms'*

```{r}
ggplot(data = surveys_complete, aes(x = weight, y = hindfoot_length)) + geom_point()
```
```{r}
surveys_plot <- ggplot(data = surveys_complete, mapping = aes( x = weight, y = hindfoot_length))

surveys_plot+
  geom_point()
```

# Challenge 1

```{r}
library(hexbin)

surveys_plot +
  geom_hex()
```
Hexagonal bin plot allows for visualization of observation clusters, but distribution of data points is much harder to notice and map to corresponding x and y axis.

# Building plots iteratively

```{r}
ggplot(data = surveys_complete, aes(x = weight, y = hindfoot_length)) + geom_point()
```

*adding transparency*

```{r}
ggplot(data = surveys_complete, aes(x = weight, y = hindfoot_length)) + geom_point(alpha = 0.1)
```

*adding color*

```{r}
ggplot(data = surveys_complete, aes(x = weight, y = hindfoot_length)) + geom_point(alpha = 0.1, color = "blue")
```

*adding color to different values*

```{r}
ggplot(data = surveys_complete, aes(x = weight, y = hindfoot_length)) + 
  geom_point(alpha = 0.1, aes(color = species_id))
```

# Challenge 2

```{r}
ggplot(data = surveys_complete, aes(x = species_id, y = weight)) +
  geom_point(aes(color = plot_type))
```

# Boxplot

```{r}
ggplot(data = surveys_complete, mapping = aes(x = species_id, y = weight)) +
    geom_boxplot()
```

*adding points*

```{r}
ggplot(data = surveys_complete, mapping = aes(x = species_id, y = weight)) +
  geom_boxplot(alpha = 0) +
    geom_jitter(alpha = 0.3, color = "tomato")
```

# Challenge 3

```{r}
ggplot(data = surveys_complete, aes(x = species_id, y = weight)) +
  geom_violin() +
    scale_y_log10()
```

*boxplot for hindfoot length*

```{r}
ggplot(surveys_complete, aes(x = species_id, y = hindfoot_length)) +
  geom_jitter(alpha = 0.5, color = "tomato") +
    geom_boxplot(alpha = 0.5, color = "sky blue")
```
