---
title: "Lab 9"
author: "Rovin Sharma"
date: "11/5/2020"
output: 
  html_document: 
    highlight: pygments
    theme: flatly
editor_options: 
  chunk_output_type: console
---

```{r, message = FALSE, error = FALSE, warning = FALSE}
library(tidyverse)
library(stringr)
```

### 14.3.1 Basic matches

Matching exact string

```{r}
x <- c("apple", "banana", "pear")
str_view(x, "an")
```

`.` matches any character

```{r}
str_view(x, ".a.")
```

`\` backslash can be used as escape to match exactly rather than use special behavior

```{r}
# To create the regular expression, we need \\
dot <- "\\."
# But the expression itself only contains one:
writeLines(dot)
```
```{r}
# And this tells R to look for an explicit .
x <- c("abc", "a.c", "bef")
str_view(x, "a\\.c")
```


To create a literal `\` you have to double it up: `\\`

```{r}
x <- "a\\b"
writeLines(x)
```
```{r}
str_view(x, "\\\\")
```

#### 14.3.1.1 Exercises

1. Explain why each of these strings don’t match a `\`: `"\"`, `"\\"`, `"\\\"`.
  
  `"\"` Will escape the next character, with no next character, it'll just keep going
  
  `"\\"` This does match `\` though, I'm not sure why it's wrong, the output shows a `\` too.
  
  `"\\\"` The first two backslashes will create a literal backslash, but then the 3rd one will escape the next character, same problem as first
  
2. How would you match the sequence `"'\`?

```{r}
  s <- "\"'\\"
  str_view(s, "\"'\\\\")
```

3. What patterns will the regular expression `\..\..\..` match? How would you represent it as a string?

It'll match a dot followed by a character 3 times

```{r}
str_view(c(".a.b.c", ".a.b", "....."), c("\\..\\..\\.."), match = TRUE)
```

### 14.3.2 Anchors

* `^` to match the start of the string.
* `$` to match the end of the string.

```{r}
x <- c("apple", "banana", "pear")
str_view(x, "^a")
```

```{r}
str_view(x, "a$")
```

Mnemonic from Evan Misshula: if you begin with power (`^`), you end up with money (`$`).

To force a regular expression to only match a complete string, anchor it with both `^` and `$`:

```{r}
x <- c("apple pie", "apple", "apple cake")
str_view(x, "apple")
```

```{r}
str_view(x, "^apple$")
```

You can also match the boundary between words with `\b`. I don’t often use this in R, but I will sometimes use it when I’m doing a search in RStudio when I want to find the name of a function that’s a component of other functions. For example, I’ll search for `\bsum\b` to avoid matching `summarise`, `summary`, `rowsum` and so on.


#### 14.3.2.1 Exercises

1. How would you match the literal string `"$^$"`?

`str_view(x, "\\$\\^\\$$")`

2. Given the corpus of common words in stringr::words, create regular expressions that find all words that:
    * Start with “y”.
    * End with “x”
    * Are exactly three letters long. (Don’t cheat by using `str_length()`!)
    * Have seven letters or more.
    
```{r}
str_view(stringr::words, "^y", match = TRUE)
str_view(stringr::words, "x$", match = TRUE)
str_view(stringr::words, "^...$", match = TRUE)
str_view(stringr::words, "^.......", match = TRUE)
```

Match argument shows only the strings that fit the criteria rather than all the strings.

### 14.3.3 Character classes and alternatives

Special patterns (like . which matches any character except for a newline):

* `\d` : matches any digit.
* `\s` : matches any whitespace (e.g. space, tab, newline).
* `[abc]` : matches a, b, or c.
* `[^abc]` : matches anything except a, b, or c.

Remember, to create a regular expression containing `\d` or `\s`, you’ll need to escape the `\` for the string, so you’ll type `"\\d"` or `"\\s"`.

A character class containing a single character is a nice alternative to backslash escapes when you want to include a single metacharacter in a regex. Many people find this more readable.

```{r}
# Look for a literal character that normally has special meaning in a regex
x <- c("abc", "a.c", "a*c", "a c")
str_view(x, "a[.]c") #surrounding . with [] targets the literal .
```

```{r}
str_view(x, ".[*]c")
```

```{r}
str_view(x, "a\\s")
# and 
str_view(x, "a[ ]")
#create the same output
```

This works for most (but not all) regex metacharacters: `$` `.` `|` `?` `*` `+` `(` `)` `[` `{`. Unfortunately, a few characters have special meaning even inside a character class and must be handled with backslash escapes: `]` `\` `^` and `-`.

You can use alternation to pick between one or more alternative patterns. For example, `abc|d..f` will match either ‘“abc”’, or "deaf". Note that the precedence for `|` is low, so that `abc|xyz` matches `abc` or `xyz` not `abcyz` or `abxyz`. Like with mathematical expressions, if precedence ever gets confusing, use parentheses to make it clear what you want:

| is basically and

```{r}
x <- c("grey", "gray")
str_view(x, "gr(e|a)y")
```


#### 14.3.3.1 Exercises

1. Create regular expressions to find all words that:
    * Start with a vowel.
    * That only contain consonants. (Hint: thinking about matching “not”-vowels.)
    * End with `ed`, but not with `eed`.
    * End with `ing` or `ise`.
```{r}
strwords <- stringr::words
str_view(strwords, "^(a|e|i|o|u)") #Vowels
str_view(strwords, "[aeiou]", match = FALSE)
str_view(strwords, "[^e]ed$", match = TRUE)
str_view(strwords, "i(ng|se)$", match = TRUE)
```
    
2. Empirically verify the rule “i before e except after c”.

```{r}
str_view(strwords, "([^c]ie|cei)", match = TRUE)
```

3. Is “q” always followed by a “u”?

```{r}
str_view(strwords, "q[^u]", match = TRUE)
```
No, it's not.

4. Write a regular expression that matches a word if it’s probably written in British English, not American English.

```{r}
str_view(strwords, "o[u]r|ise$", match = TRUE)
```

5. Create a regular expression that will match telephone numbers as commonly written in your country.

```{r}
str_view(strwords, "\\d\\d\\d-\\d\\d\\d-\\d\\d\\d\\d")
```

### 14.3.4 Repetition

The next step up in power involves controlling how many times a pattern matches:

* `?`: 0 or 1
* `+`: 1 or more
* `*`: 0 or more

```{r}
x <- "1888 is the longest year in Roman numerals: MDCCCLXXXVIII"
str_view(x, "CC?")
```

```{r}
str_view(x, "C+")
```

```{r}
str_view(x, 'C[LX]+')
```

Note that the precedence of these operators is high, so you can write: `colou?r` to match either American or British spellings. That means most uses will need parentheses, like `bana(na)+`.

You can also specify the number of matches precisely:

* `{n}`: exactly n
* `{n,}`: n or more
* `{,m}`: at most m
* `{n,m}`: between n and m

```{r}
str_view(x, "C{2}")
```
```{r}
str_view(x, "C{2,}")
```
```{r}
str_view(x, "C{2,}")
```

By default these matches are “greedy”: they will match the longest string possible. You can make them “lazy”, matching the shortest string possible by putting a `?` after them. This is an advanced feature of regular expressions, but it’s useful to know that it exists:

```{r}
str_view(x, 'C{2,3}?')
```
```{r}
str_view(x, 'C[LX]+?')
```


#### 14.3.4.1 Exercises

1. Describe the equivalents of ?, +, * in {m,n} form.

```{r}
str_view(x, 'C{0,1}')
```
```{r}
str_view(x, 'C{1,}')
```
```{r}
str_view(x, 'C{0,}')
```
2. Describe in words what these regular expressions match: (read carefully to see if I’m using a regular expression or a string that defines a regular expression.)
    * `^.*$` Matches any string 
    * `"\\{.+\\}"` Anything with curly braces
    * `\d{4}-\d{2}-\d{2}` Four digits with hypens b/t
    * `"\\\\{4}"` Matches 4 \
3. Create regular expressions to find all words that:
    * Start with three consonants.
    str_view(words, "^[^aeiou]{3}", match = TRUE)
    * Have three or more vowels in a row.
    str_view(words, "[aeiou]{3,}", match = TRUE)
    * Have two or more vowel-consonant pairs in a row.
    str_view(words, "([aeiou][^aeiou]){2,}", match = TRUE)
    

### 14.3.5 Grouping and backreferences

Earlier, you learned about parentheses as a way to disambiguate complex expressions. Parentheses also create a numbered capturing group (number 1, 2 etc.). A capturing group stores the part of the string matched by the part of the regular expression inside the parentheses. You can refer to the same text as previously matched by a capturing group with backreferences, like `\1`, `\2` etc. For example, the following regular expression finds all fruits that have a repeated pair of letters.

```{r}
str_view(fruit, "(..)\\1", match = TRUE)
```

(Shortly, you’ll also see how they’re useful in conjunction with `str_match()`.)


#### 14.3.5.1 Exercises

1. Describe, in words, what these expressions will match:
    * `(.)\1\1`
    One character repeated twice more
    * `"(.)(.)\\2\\1"`
    Two characters repeated once with flipped position
    * `(..)\1`
    Two characters repeated once
    * `"(.).\\1.\\1"`
    One character followed by any character, repeated with another character between
    * `"(.)(.)(.).*\\3\\2\\1"`
    3 characters, 0 or more original character, repeated original 3 characters in reverse order
2. Construct regular expressions to match words that:
    * Start and end with the same character.
    (x, "^(.)((.*\\1$)|\\1?$)")
    * Contain a repeated pair of letters (e.g. “church” contains “ch” repeated twice.)
    "([A-Za-z][A-Za-z]).*\\1"
    * Contain one letter repeated in at least three places (e.g. “eleven” contains three “e”s.)
    "([a-z]).*\\1.*\\1"

## 14.4 Tools

Now that you’ve learned the basics of regular expressions, it’s time to learn how to apply them to real problems. In this section you’ll learn a wide array of stringr functions that let you:

* Determine which strings match a pattern.
* Find the positions of matches.
* Extract the content of matches.
* Replace matches with new values.
* Split a string based on a match.

### 14.4.1 Detect matches

* Note the below takes advantage of the built in `word` vector that is loaded with the `stringr` library.

To determine if a character vector matches a pattern, use str_detect(). It returns a logical vector the same length as the input:

```{r}
x <- c("apple", "banana", "pear")
sum(str_detect(x, "e"))
```

Remember that when you use a logical vector in a numeric context, `FALSE` becomes 0 and `TRUE` becomes 1. That makes `sum()` and `mean()` useful if you want to answer questions about matches across a larger vector:

```{r}
# How many common words start with t?
sum(str_detect(words, "^t"))
```

```{r}
# What proportion of common words end with a vowel?
mean(str_detect(words, "[aeiou]$"))
```

When you have complex logical conditions (e.g. match a or b but not c unless d) it’s often easier to combine multiple `str_detect()` calls with logical operators, rather than trying to create a single regular expression. For example, here are two ways to find all words that don’t contain any vowels:

```{r}
# Find all words containing at least one vowel, and negate
no_vowels_1 <- !str_detect(words, "[aeiou]")
```
```{r}
# Find all words consisting only of consonants (non-vowels)
no_vowels_2 <- str_detect(words, "^[^aeiou]+$")
identical(no_vowels_1, no_vowels_2)
```

The results are identical, but I think the first approach is significantly easier to understand. If your regular expression gets overly complicated, try breaking it up into smaller pieces, giving each piece a name, and then combining the pieces with logical operations.

A common use of `str_detect()` is to select the elements that match a pattern. You can do this with logical subsetting, or the convenient `str_subset()` wrapper:

```{r}
words[str_detect(words, "x$")]
```

```{r}
str_subset(words, "x$")
```

Typically, however, your strings will be one column of a data frame, and you’ll want to use filter instead:

```{r}
df <- tibble(
  word = words, 
  i = seq_along(word)
)
df %>% 
  filter(str_detect(word, "x$"))
```

A variation on `str_detect()` is `str_count()`: rather than a simple yes or no, it tells you how many matches there are in a string:

```{r}
x <- c("apple", "banana", "pear")
str_count(x, "a")
```
```{r}
# On average, how many vowels per word?
mean(str_count(words, "[aeiou]"))
```

It’s natural to use `str_count()` with `mutate()`:
```{r}
df %>% 
  mutate(
    vowels = str_count(word, "[aeiou]"),
    consonants = str_count(word, "[^aeiou]")
  )
```

Note that matches never overlap. For example, in `"abababa"`, how many times will the pattern `"aba"` match? Regular expressions say two, not three:

```{r}
str_count("abababa", "aba")
```
```{r}
str_view_all("abababa", "aba")
```

Note the use of `str_view_all()`. As you’ll shortly learn, many stringr functions come in pairs: one function works with a single match, and the other works with all matches. The second function will have the suffix `_all`.


#### 14.4.1.1 Exercises

1. For each of the following challenges, try solving it by using both a single regular expression, and a combination of multiple str_detect() calls.
    * Find all words that start or end with x.
    ```{r}
    str_detect(words, "^x|x$")
    ```
    * Find all words that start with a vowel and end with a consonant.
    ```{r}
    str_subset(words, "^[aeiou].*[^aeiou]$")
    ```

2. What word has the highest number of vowels? What word has the highest proportion of vowels? (Hint: what is the denominator?)
```{r}
vowels <- str_count(words, "[aeiou]")
words[which(vowels == max(vowels))]
```

```{r}
prop_vowels <- str_count(words, "[aeiou]") / str_length(words)
words[which(prop_vowels == max(prop_vowels))]
```

### 14.4.2 Extract matches

To extract the actual text of a match, use `str_extract()`. To show that off, we’re going to need a more complicated example. I’m going to use the Harvard sentences, which were designed to test VOIP systems, but are also useful for practicing regexps. These are provided in `stringr::sentences`:

```{r}
length(sentences)
```
```{r}
head(sentences)
```

Imagine we want to find all sentences that contain a colour. We first create a vector of colour names, and then turn it into a single regular expression:

```{r}
colours <- c("red", "orange", "yellow", "green", "blue", "purple")
colour_match <- str_c(colours, collapse = "|")
colour_match
```

Now we can select the sentences that contain a colour, and then extract the colour to figure out which one it is:

```{r}
has_colour <- str_subset(sentences, colour_match)
matches <- str_extract(has_colour, colour_match)
head(matches)
```
* The above example has an error. 
```{r}
has_colour
```
Can your spot the issue with the regular expression?

Note that `str_extract()` only extracts the first match. We can see that most easily by first selecting all the sentences that have more than 1 match:

```{r}
more <- sentences[str_count(sentences, colour_match) > 1]
str_view_all(more, colour_match)
```
```{r}
str_extract(more, colour_match)
```

This is a common pattern for stringr functions, because working with a single match allows you to use much simpler data structures. To get all matches, use `str_extract_all()`. It returns a list:
```{r}
str_extract_all(more, colour_match)
```
You’ll learn more about lists in [lists](https://r4ds.had.co.nz/vectors.html#lists) and [iteration](https://r4ds.had.co.nz/vectors.html#iteration).

If you use `simplify = TRUE`, `str_extract_all()` will return a matrix with short matches expanded to the same length as the longest:

```{r}
str_extract_all(more, colour_match, simplify = TRUE)
```
```{r}
x <- c("a", "a b", "a b c")
str_extract_all(x, "[a-z]", simplify = TRUE)
```

####  14.4.2.1 Exercises

1. In the previous example, you might have noticed that the regular expression matched “flickered”, which is not a colour. Modify the regex to fix the problem.
colour_match has to be changed to add a \b before and after pattern for space: str_c("\\b(", str_c(colours, collapse = "|"), ")\\b")
2. From the Harvard sentences data, extract:
    * The first word from each sentence.
    str_extract(sentences, "[A-Za-z][A-Za-z']*")
    * All words ending in ing.
    pattern <- "\\b[A-Za-z]+ing\\b"
sentences_with_ing <- str_detect(sentences, pattern)
unique(unlist(str_extract_all(sentences[sentences_with_ing], pattern)))
    str_extract(sentences, ")
    * All plurals.
    unique(unlist(str_extract_all(sentences, "\\b[A-Za-z]{3,}s\\b")))

### 14.4.3 Grouped matches

Earlier in this chapter we talked about the use of parentheses for clarifying precedence and for backreferences when matching. You can also use parentheses to extract parts of a complex match. For example, imagine we want to extract nouns from the sentences. As a heuristic, we’ll look for any word that comes after “a” or “the”. Defining a “word” in a regular expression is a little tricky, so here I use a simple approximation: a sequence of at least one character that isn’t a space.

```{r}
noun <- "(a|the) ([^ ]+)"
has_noun <- sentences %>%
  str_subset(noun) %>%
  head(10)
has_noun %>% 
  str_extract(noun)
```
`str_extract()` gives us the complete match; `str_match()` gives each individual component. Instead of a character vector, it returns a matrix, with one column for the complete match followed by one column for each group:

```{r}
has_noun %>% 
  str_match(noun)
```

(Unsurprisingly, our heuristic for detecting nouns is poor, and also picks up adjectives like smooth and parked.)

If your data is in a tibble, it’s often easier to use `tidyr::extract()`. It works like `str_match()` but requires you to name the matches, which are then placed in new columns:

```{r}
tibble(sentence = sentences) %>% 
  tidyr::extract(
    sentence, c("article", "noun"), "(a|the) ([^ ]+)", 
    remove = FALSE
  )
```

Like `str_extract()`, if you want all matches for each string, you’ll need `str_match_all()`.

#### 14.4.3.1 Exercises

1. Find all words that come after a “number” like “one”, “two”, “three” etc. Pull out both the number and the word.
```{r}
numword <- "\\b(one|two|three|four|five|six|seven|eight|nine|ten) +(\\w+)"
sentences[str_detect(sentences, numword)] %>%
  str_extract(numword)
```


### 14.4.4 Replacing matches

`str_replace()` and `str_replace_all()` allow you to replace matches with new strings. The simplest use is to replace a pattern with a fixed string:

```{r}
x <- c("apple", "pear", "banana")
str_replace(x, "[aeiou]", "-")
```
```{r}
str_replace_all(x, "[aeiou]", "-")
```

With `str_replace_all()` you can perform multiple replacements by supplying a named vector:

```{r}
x <- c("1 house", "2 cars", "3 people")
str_replace_all(x, c("1" = "one", "2" = "two", "3" = "three"))
```

Instead of replacing with a fixed string you can use backreferences to insert components of the match. In the following code, I flip the order of the second and third words.

```{r}
sentences %>% 
  str_replace("([^ ]+) ([^ ]+) ([^ ]+)", "\\1 \\3 \\2") %>% 
  head(5)
```


#### 14.4.4.1 Exercises

1. Replace all forward slashes in a string with backslashes.
str_replace_all("past/present/future", "/", "\\\\")
2. Implement a simple version of str_to_lower() using replace_all().
```{r}
replacements <- c("A" = "a", "B" = "b", "C" = "c", "D" = "d", "E" = "e",
                  "F" = "f", "G" = "g", "H" = "h", "I" = "i", "J" = "j", 
                  "K" = "k", "L" = "l", "M" = "m", "N" = "n", "O" = "o", 
                  "P" = "p", "Q" = "q", "R" = "r", "S" = "s", "T" = "t", 
                  "U" = "u", "V" = "v", "W" = "w", "X" = "x", "Y" = "y", 
                  "Z" = "z")
lower_words <- str_replace_all(words, pattern = replacements)
head(lower_words)
```

3. Switch the first and last letters in words. Which of those strings are still words?

```{r}
swapped <- str_replace_all(words, "^([[:alpha:]])(.*)([[:alpha:]])$", "\\3\\2\\1")
intersect(swapped, words)
```

### 14.4.5 Splitting

Use `str_split()` to split a string up into pieces. For example, we could split sentences into words:

```{r}
sentences %>%
  head(5) %>% 
  str_split(" ")
```
Because each component might contain a different number of pieces, this returns a list. If you’re working with a length-1 vector, the easiest thing is to just extract the first element of the list:
```{r}
"a|b|c|d" %>% 
  str_split("\\|") %>% 
  .[[1]]
```
Otherwise, like the other stringr functions that return a list, you can use `simplify = TRUE` to return a matrix:
```{r}
sentences %>%
  head(5) %>% 
  str_split(" ", simplify = TRUE)
```
You can also request a maximum number of pieces:

```{r}
fields <- c("Name: Hadley", "Country: NZ", "Age: 35")
fields %>% str_split(": ", n = 2, simplify = TRUE)
```

Instead of splitting up strings by patterns, you can also split up by character, line, sentence and word `boundary()`s:
```{r}
x <- "This is a sentence.  This is another sentence."
str_view_all(x, boundary("word"))
```
```{r}
str_split(x, " ")[[1]]
```
```{r}
str_split(x, boundary("word"))[[1]]
```

#### 14.4.5.1 Exercises

1. Split up a string like "apples, pears, and bananas" into individual components.

```{r}
str_split(words, ", +(and +)?")[[1]]
```

2. Why is it better to split up by boundary("word") than " "?

Space will retain punctuation

3. What does splitting with an empty string ("") do? Experiment, and then read the documentation.

```{r}
str_split("Hello my name is Rovin. I like to eat food to survive!", "")[[1]]
```
It splits every single character, including white spaces.
