---
title: "Untitled"
author: "Rovin Sharma"
date: "11/30/2020"
output: html_document
---

## Control Structures

Commonly used control structures are

- `if` and `else`: testing a condition and acting on it

- `for`: execute a loop a fixed number of times 

- `while`: execute a loop _while_ a condition is true 

### Conditional statements with `if`-`else`

The `if`-`else` combination is probably the most commonly used control structure in R (or perhaps any language). This structure allows you to test a condition and act on it depending on whether it's true or false. 

Here is an example of a valid if/else structure.

```{r,prompt=FALSE}
## Generate a uniform random number
x <- runif(1, 0, 10)  
if(x > 3) {
        y <- 10
} else {
        y <- 0
}
```

The value of `y` is set depending on whether `x > 3` or not. 

### Important Operators

#### Comparison operators

* `==` (equal)
* `!=` (not equal)
* `>` (greater than)
* `>=` (greater than or equal)
* `<` (less than)
* `<=` (less than or equal)

#### Logical operators
		
* `&` (and)
* `|` (or) 
* `!` (not)

### `for` Loops

In R, for loops take an interator variable and assign it successive values from a sequence or vector. For loops are most commonly used for iterating over the elements of an object (list, vector, etc.)

```{r}
for(i in 1:10) {
        print(i)
}

for(i in 1:10) {
  if(i == 1) {print("a")}
  else{print("b")}
}
```

This loop takes the `i` variable and in each iteration of the loop gives it values 1, 2, 3, ..., 10, executes the code within the curly braces, and then the loop exits.

The following three loops all have the same behavior.

```{r}
x <- c("a", "b", "c", "d")
for(i in 1:5) {
        ## Print out each element of 'x'
        print(x[i])  
}
```

The `seq_along()` function is commonly used in conjunction with for loops in order to generate an integer sequence based on the length of an object (in this case, the object `x`).

```{r}
## Generate a sequence based on length of 'x'
for(i in seq_along(x)) {   
        print(x[i])
}
```

It is not necessary to use an index-type variable. 

```{r}
for(letter in x) {
        print(letter)
}
```

For one line loops, the curly braces are not strictly necessary.

```{r}
for(i in 1:4) print(x[i])
```

### Nested `for` loops

`for` loops can be nested inside of each other.

```r
x <- matrix(1:6, 2, 3)
for(i in seq_len(nrow(x))) {
        for(j in seq_len(ncol(x))) {
                print(x[i, j])
        }   
}
```

Nested loops are commonly needed for multidimensional or hierarchical data structures (e.g. matrices, lists). Be careful with nesting though. Nesting beyond 2 to 3 levels often makes it difficult to read/understand the code. If you find yourself in need of a large number of nested loops, you may want to break up the loops by using functions (discussed later).

### `while` Loops

While loops begin by testing a condition. If it is true, then they execute the loop body. Once the loop body is executed, the condition is tested again, and so forth, until the condition is false, after which the loop exits.

```{r}
count <- 0
while(count < 10) {
        print(count)
        count <- count + 1
}
```

While loops can potentially result in infinite loops if not written properly. Use with care!

Sometimes there will be more than one condition in the test.

```{r}
z <- 5
set.seed(1)
while(z >= 3 && z <= 10) {
        coin <- rbinom(1, 1, 0.5)
        
        if(coin == 1) {  ## random walk
                z <- z + 1
        } else {
                z <- z - 5
        } 
}
print(z)
```

Conditions are always evaluated from left to right. 

### Functions in R

Functions in R are "first class objects", which means that they can be treated much like any other R object. Importantly,

- Functions can be passed as arguments to other functions. This is very handy for the various apply functions, like `lapply()` and `sapply()` that we will learn below.

- Functions can be nested, so that you can define a function inside of
  another function

### Your First Function

Functions are defined using the `function()` directive and are stored
as R objects just like anything else. In particular, they are R
objects of class "function".

Here's a simple function that takes no arguments and does nothing.

```{r}
f <- function() {
        ## This is an empty function
}
## Functions have their own class
class(f)  
## Execute this function
f()       
```

The next thing we can do is create a function that actually has a non-trivial *function body*.

```{r}
f <- function() {
        cat("Hello, world!")
}
f()
```

The last aspect of a basic function is the *function arguments*. These are the options that you can specify to the user that the user may explicity set. For this basic function, we can add an argument that determines how many times "Hello, world!" is printed to the console.

```{r}
f <- function(num) {
        for(i in seq_len(num)) {
                cat("Hello, world!\n") ## the \n creates a new line, otherwise it would print Hello, world!Hello, world!Hello, world!
        }
}
f(3)
```

Obviously, we could have just cut-and-pasted the `cat("Hello, world!\n")` code three times to achieve the \ same effect, but then we wouldn't be programming, would we? Also, it would be un-neighborly of you to give your code to someone else and force them to cut-and-paste the code however many times the need to see "Hello, world!".


```{r}
f <- function(num) {
        hello <- "Hello, world!\n"
        for(i in seq_len(num)) {
                cat(hello)
        }
        chars <- nchar(hello) * num
        chars
}
meaningoflife <- f(1)
print(meaningoflife)
```

In the above function, we didn't have to indicate anything special in order for the function to return the number of characters. In R, the return value of a function is always the very last expression that is evaluated. Because the `chars` variable is the last expression that is evaluated in this function, that becomes the return value of the function.

Note that there is a `return()` function that can be used to return an explicity value from a function, but it is rarely used in R (we will discuss it a bit later in this chapter).

the num in f(num) must be defined, so the below e.g. doesn't work.

```{r, eval=FALSE}
f()
```

We can modify this behavior by setting a *default value* for the argument `num`. Any function argument can have a default value, if you wish to specify it. Sometimes, argument values are rarely modified (except in special cases) and it makes sense to set a default value for that argument. This relieves the user from having to specify the value of that argument every single time the function is called. 

Here, for example, we could set the default value for `num` to be 1, so that if the function is called without the `num` argument being explicitly specified, then it will print "Hello, world!" to the console once.

```{r}
f <- function(num = 5) {
        hello <- "Hello, world!\n"
        for(i in seq_len(num)) {
                cat(hello)
        }
        chars <- nchar(hello) * num
        chars
}
f()    ## Use default value for 'num'
f(2)   ## Use user-specified value
```

Remember that the function still returns the number of characters printed to the console.

At this point, we have written a function that 

* has one *formal argument* named `num` with a *default value* of 1. The _formal arguments_ are the arguments included in the function definition. The `formals()` function returns a list of all the formal arguments of a function 

* prints the message "Hello, world!" to the console a number of times indicated by the argument `num`

* *returns* the number of characters printed to the console


Functions have _named arguments_ which can optionally have default values. Because all function arguments have names, they can be specified using their name.

```{r}
z <- function(num = 6) {
  while(num < 5) {
    cat("Less than 5")
  }
}
z()
```
